---
title: "Replication of Study 2 by Wiltermuth (2011)"
author: "Em Reit (emreit@stanford.edu)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
---


##Introduction
For my replication project, I chose Wiltermuth's 2011 paper, *Cheating More When the Spoils are Split.* The reason I chose Wiltermuth's 2011 paper is because I am working on some research about cheating behavior, and we were hoping to actually use the exact cheating paradigm that he used in his Study 2. The broad point of his paper was to demonstrate that people are more likely to engage in cheating behavior when they are not the sole person benfiting from such behavior. In his Study 2, participants were given an anagram unscrambling task that included a nearly unsolvable word, thus giving them the opportunity to (falsely) report their performance. He manipulated if people thought they would be the sole recipient of some sort of bonus pay-out per word unscrambled, or if someone else would also benefit from their performance. He found that people who thought someone else would also benefit from the unscrambling of the words were more likely to cheat, as measured by the reporting of having unscrambled the third (unsolvable) word.



##Methods

###Power Analysis

The original effect size of the main test (chi-square between those who cheat v. not in the self alone as compared to the self-and-other) was .14. A power analysis revealed that we would need a sample of 401 participants to detect such an effect with 80% power. Given bugdet constraints of having $50 to run the study, and factoring each participant would get paid 40 cents base pay + some variable pay, which is on average 90 cents total per participant, we intend to collect 40 participant data.

###Planned Sample

Our planned sample size is 40 participants. Given we are more interested in how the tasks functions on MTurk, and due to budgetary constraints of the course, we will stick with a small sample and not generalize about the validity of the finding because it will not be well powered.

###Materials

We were provided with the following materials: 

"Please unscramble each of the word jumbles below. Remember you will be only be paid for those you solve in order from the beginning.&nbsp; After you unscramble a word jumble, please check Solved to indicate that you have solved the word jumble

Keep in mind that you will receive $1 for each word jumble you solve in order from the beginning.  Also, remember that another randomly selected participant from another experiment will also receive $1 for each word jumble you solve in order from the beginning.

Please click continue after you have finished or nine minutes have elapsed.

Thank you

Anagram1	likert	1. U N H T E D	2	Not Solved	Solved


Anagram2	likert	2. E O S H U	2	Not Solved	Solved


Anagram3	likert	3. U N A A G T	2	Not Solved	Solved


Anagram4	likert	4. Y T H O I R D	2	Not Solved	Solved


Anagram5	likert	5. O L A R C	2	Not Solved	Solved


Anagram6	likert	6. J N I P M U G	2	Not Solved	Solved


Anagram7	likert	7. H G I T W E	2	Not Solved	Solved


Anagram8	likert	8. C L A L S O U	2	Not Solved	Solved


Anagram9	likert	9. Y O M S E E V L D	2	Not Solved	Solved"


Participants will see nine word jumbles, which are displayed above. Unbeknownst to the participants, the third word jumble could only be unscrambled to spell the obscure word "taguan." In Wiltermuth's pre-testing, none of the thirty participants successfully unscrambled this word jumble. Thus, it would be unlikely that participants acting honestly would report having solved this jumble.


###Procedure	
We selected **two** of the five conditions that Wiltermuth ran for the purposes of this replication:

Self-and-other benefit: You will receive $1 per word jumble solved. Additionally, a randomly-selected participant from another experiment will also receive $1 per word jumble solved.

Self-alone high benefit: You will receive $2 per word jumble solved.

Participants were either shown the self-and-other benefit condition, or the self-alone high benefit condition. **Due to budget constraints, in the self-and-other benefit condition, the participant and the randomly-selected participant received 5 cents, and in the self-alone condition, the participant received 10 cents.*

###Analysis Plan

Our key analysis is a chi-square test of the number of people who cheat v. don't cheat in each of the two conditions. We will not exclude any participants in our main analysis. 

Additional analyses will include exclusion based on (1) if the participant states having ran into any issues, and (2) if the participant states having done a task similar to the one we asked them to do.

###Differences from Original Study

There are a few major differences between this replication and the original study. First, Wiltermuth used an undergraduate online sample, whereas we are using the MTurk population. As his sample was less familiar with online studies (and have a much lower likelihood of having encountered this exact paradigm before), they might reflect a lower cheating rate, overall.

Second, our pay-off structure - though it is the same ratio - is very different. Wiltermuth paid $1 per additional word unscrambled in self-and-other condition, and $2 in self-alone condition. We are paying 5 cents in self-and-other, and 10 cents in self alone. In this way, we maintaing that the self-alone receives twice as much as the self-and-other, but it is significantly less money than in the original. This could lead to participants taking cheating less seriously, and overreporting it in general. 

Lastly, we include some possible exclusion criteria for exploratory purposes, whereas Wiltermuth mention no exclusions. 

### Methods Addendum (Post Data Collection)

You can comment this section out prior to final report with data collection.

#### Actual Sample
  Sample size, demographics, data exclusions based on rules spelled out in analysis plan

#### Differences from pre-data collection methods plan
  Any differences from what was described as the original plan, or “none”.


##Results


### Data preparation

Data preparation following the analysis plan.
	

```{r}
library(car)
library(psych)
library(tidyverse)
```


```{r}
#Load in raw data.
setwd("~/Documents/StanfordCoursework/winter2017/MF's/Project/wiltermuth2011/Pilot B")

#Remove the first two header lines from Qualtrics raw file.

qualtrics <- read.csv(("wiltermuth2011_pilotB.csv"), header = TRUE, stringsAsFactors = FALSE) %>%
  slice(-(1:2))
```


```{r cheating count}
d <- qualtrics

#Embedded data is set to Condition = self or other. These are the two conditions used for analyses.
table(d$Condition)

#self_cheat and other_cheat are the binary "yes/no" variables about having solved the unsovlable 3rd word.
d$self_cheat <- as.factor(d$self_cheat)
d$other_cheat <- as.factor(d$other_cheat)

#this combines the cheating measure across conditions for analyses.
d1 <- d %>%
  mutate(cheat_yes = self_cheat == 2 | other_cheat == 2)
table(d1$cheat_yes) 
```


```{r frequency of cheating per condition}
#this breaks down cheating (false/true) by condition (self/other)
cheatfrequency <- d1 %>% 
  group_by(Condition, cheat_yes) %>% 
  dplyr::summarise(freq=n())
print(cheatfrequency)

cheatyes_self_freq <- 1
cheatno_self_freq <- 1
cheatyes_other_freq <- 2
cheatno_other_freq <- 0
```


```{r chi-square test of interest}
#this tests the above frequencies to see if the % who cheat in self-alone is different than self-other
prop <- as.table(matrix(c(cheatyes_self_freq, cheatno_self_freq, cheatyes_other_freq, cheatno_other_freq), byrow=T,ncol=2))
print(prop)
chisq.test(prop)


#setting up values to graph
percent_cheat_self <- (cheatyes_self_freq/(cheatyes_self_freq+cheatno_self_freq))*100; percent_cheat_self
percent_cheat_other <- (cheatyes_other_freq/(cheatyes_other_freq+cheatno_other_freq))*100; percent_cheat_other


#this is the original % of cheating in self-alone vs. self-other in Wiltermuth 2011
originalprop <- as.table(matrix(c(15,52,22,38), byrow=T, ncol=2))
print(originalprop)
chisq.test(originalprop)
percent_cheat_self_original <- (15/67)*100
percent_cheat_other_original <- (22/60)*100

library(ggplot2)

counts <- as.data.frame(matrix(c(percent_cheat_self, percent_cheat_other, percent_cheat_self_original, percent_cheat_other_original, "replication", "replication", "original", "original", "self", "other", "self", "other"), byrow=F, ncol=3))
counts$V1 <- as.numeric(as.character(counts$V1))
counts$other_v_self <- counts$V3
counts


ggplot(data = counts, aes(x = V2, y = V1, group = other_v_self, fill = other_v_self)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Percent Cheat for Self v. Self-And-Other: Original Versus Replication", x = "Original v. Replication", y = "% Cheat")
```




###Exploratory analyses

Any follow-up analyses desired (not required).  

## Discussion

### Summary of Replication Attempt

Open the discussion section with a paragraph summarizing the primary result from the confirmatory analysis and the assessment of whether it replicated, partially replicated, or failed to replicate the original result.  

### Commentary

Add open-ended commentary (if any) reflecting (a) insights from follow-up exploratory analysis, (b) assessment of the meaning of the replication (or not) - e.g., for a failure to replicate, are the differences between original and present study ones that definitely, plausibly, or are unlikely to have been moderators of the result, and (c) discussion of any objections or challenges raised by the current and original authors about the replication attempt.  None of these need to be long.
